Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-02-23 paper/main.tex
small progress in Section 3

These observations motivate our study of a technique reducing the
number of acceptance marks without altering the structure of the
automaton. This paper presents a technique heavily based on
\emph{quantified Boolean formulas (QBF)}. For a given TELA and
parameters $c,m$, it produces a QBF formula which is satisfiable if
and only if there exists an automaton with the same structure, $m$
acceptance marks, an acceptance formula in disjunctive normal form
with $c$ clauses, and the same set of accepting runs as the original
automaton. The placement of the marks on transitions and the
acceptance formula can be obtained from a model of the
formula. Besides this formula, we describe also the construction of
two simpler formulas such that their satisfiability implies the
existence of an automaton with the same structure, $m$ acceptance
marks, and the same set of accepting runs, but not vice versa.

@@

\emph{Quantified Boolean formulas (QBFs)} are Boolean formulas
extended with universal and existential quantification over
propositional variables. We assume that subformulas of the form
$\forall x.\psi$ and $\exists x.\psi$ do not contain another
quantification of variable $x$ inside $\psi$. The semantics of
$\forall x.\psi$ and $\exists x.\psi$ is given by equivalences
\[
  \begin{array}{rcl}
    \forall x.\psi &~\equiv~& \psi[x\rightarrow\true]~\wedge~\psi[x\rightarrow\false]\\
    \exists x.\psi &~\equiv~& \psi[x\rightarrow\true]~\vee~   \psi[x\rightarrow\false]
  \end{array}
\]
where $\psi[x\rightarrow \rho]$ denotes the formula $\psi$ with all
occurrences of $x$ simultaneously replaced by $\rho$. The equivalences
imply that each QBF can be transformed into an equivalent Boolean
formula. However, the size of this Boolean formula can be exponential
in the size of the original QBF. Let $V$ be the set of all
propositional variables. A mapping $\mu:V\rightarrow\{\true,\false\}$
is a \emph{model} of a QBF formula $\varphi$ iff it is a satisfying
assignment of the equivalent Boolean formula. A QBF formula is
\emph{satisfiable} iff it has a model.

@@

@@

@@

>>>>

These observations motivate our study of a technique reducing the
number of acceptance marks without altering the structure of the
automaton. This paper presents a technique heavily based on
\emph{quantified Boolean formulas (QBF)}. For a given TELA and
parameters $C,K$, it produces a QBF formula which is satisfiable if
and only if there exists an automaton with the same structure, $K$
acceptance marks, an acceptance formula in disjunctive normal form
with $C$ clauses, and the same set of accepting runs as the original
automaton. The placement of the marks on transitions and the
acceptance formula can be obtained from a model of the
formula. Besides this formula, we describe also the construction of
two simpler formulas such that their satisfiability implies the
existence of an automaton with the same structure, $K$ acceptance
marks, and the same set of accepting runs, but not vice versa.

@@

\emph{Quantified Boolean formulas (QBF)} are Boolean formulas
extended with universal and existential quantification over
propositional variables. We assume that subformulas of the form
$\forall x.\psi$ and $\exists x.\psi$ do not contain another
quantification of variable $x$ inside $\psi$. The semantics of
$\forall x.\psi$ and $\exists x.\psi$ is given by equivalences
\[
  \begin{array}{rcl}
    \forall x.\psi &~\equiv~& \psi[x\rightarrow\true]~\wedge~\psi[x\rightarrow\false]\\
    \exists x.\psi &~\equiv~& \psi[x\rightarrow\true]~\vee~   \psi[x\rightarrow\false]
  \end{array}
\]
where $\psi[x\rightarrow \rho]$ denotes the formula $\psi$ with all
occurrences of $x$ simultaneously replaced by $\rho$. The equivalences
imply that each QBF can be transformed into an equivalent Boolean
formula. However, the size of this Boolean formula can be exponential
in the size of the original QBF. Let $V$ be the set of all
propositional variables. A mapping $\mu:V\rightarrow\{\true,\false\}$
is a \emph{model} of a QBF formula $\varphi$ iff it is a satisfying
assignment of the equivalent Boolean formula. A QBF formula is
\emph{satisfiable} iff it has a model.

@@

Recall that we aim to reduce the number of acceptance marks in a given
TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ without altering its
structure and language. In other words, we change only the acceptance
marks on individual transitions and we replace the acceptance formula
by a new one with less acceptance marks. To ensure that the modified
automaton is equivalent to the original one, we actually construct the
new acceptance formula such that the set of acceptance runs (modulo
changed acceptance marks on transitions) is preserved.

We start with the following simple observation. Definovat $inf(\varphi)$?

\textbf{Jan is here!}

Recall that our goal is to reduce the number of acceptance marks in a
given TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ without altering its
structure and language. Hence, our reduction replaces the set $M$ of
marks by a smaller set, the acceptance formula $\varphi$ by a new one,
and changes the sets of acceptance marks on individual transitions.

Let us assume that we are given a TELA $\mA$ and two integer
parameters $C,K\ge 0$, where $K$ is the desired number of acceptance
marks after the reduction and $C$ is the desired number of clauses in
the new acceptance formula in DNF. This section describes the
construction of a QBF formula for $\mA,C,K$ such that satisfiability
of this formula implies existence of the reduced automata with

This section explains the construction of QBF formulas for a given
TELA $\mA =(Q,M,\Sigma,q_I,\delta,\varphi)$ and two integer parameters
$C,K\ge 0$ such that satisfiability of such a formula implies that
there exists an acceptance formula in DNF with $C$ clauses and $K$
acceptance marks and a placement of these marks on automata
transitions such that $\mA$ with this acceptance formula and
acceptance marks and acceptance formula

such that their model describes an acceptance formula in DNF
with $C$ clauses and $K$ acceptance marks and placement of these marks
on automata transitions such that

@@

Pridat do experimentu nejaky graf, ktery ukaze, jak rychle se nektere
automaty redukovaly na minimum? Nebo spis ukazat vliv timeoutu na
redukce?  (zvysovat timeout od 1s do 50s a vzdy ukazat miru
redukce). Ty redukce ukazovat nejak jinak nez jen celkovym cislem
(boxplot? Scatterplot?).

Zkusit upravit redukci tak, aby vysledkem bylo TGBA a zkusit dale
redukovat akceptacni znacky TGBA automatu redukovanych Spotem?

Chceme ukazovat, ze redukce znacek muze nekdy umoznit dalsi redukci stavu?

@@

Redukce netrva vecne, da se prakicky pouzit a zkracenim timeoutu na
QBF-solver se da zrychlit (tradeoff s ucinnosti). Prinejmensim u
Rabinizeru 4 je velmi dobra sance, ze se to vyplati. Pridat do
experimentu nejaky graf, ktery ukaze, jak rychle se nektere automaty
redukovaly na minimum?
